Augmenting the DonorsChoose.org Corpus for Meta-Learning

### 2019-03-31 doc/paper/main.tex
paper/main.tex: Shorten and tweak choice of words

\section{Introduction}
Meta-Learning is the process of applying machine learning algorithms on metadata as generated by a subsystem composed of machine learners. The goal of meta-learning systems is to improve existing approaches by combining the strengths of several single machine learning systems~\cite{AIR:MetalearningSurveyOfTrendsAndTechnologiesLemke2015}. In recommender systems it is an emerging field of interest as meta-learning promises to better sort through item-, user- and transaction-information as to retrieve the best possible match of item and user. In order to investigate meta-learning algorithms, the required metadata needs to be computed in advance.

One challenge for researchers interested in meta-learning is that suitable datasets are time-consuming and cumbersome to construct. Creating a system of individual machine learning experiments and amending their metadata to the original corpus takes time and resources away from working with the actual meta-learners. Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}. Furthermore, a noteworthy shortcoming of the metadata constructed in the aforementioned publications is the usage of a metric which limits the choice of algorithms. The RMSE of the predicted and actual score is a popular metric but, e.g., can not be computed in a straight-forward way for content-based filtering algorithms.

Our objective is to address the data augmentation part in a comprehensive and reproducible~\cite{UMUAI:TowardsReproducibilityInRecSysBeel2016} way as to allow researchers to focus on the actual meta-learning evaluation. We are presenting a public domain dataset which to our best knowledge has not been used for meta-learning before in the scientific community, yet due to its feature-rich, vast corpus is a suitable candidate. The raw dataset is augmented with metadata of several machine learning algorithms from a variety of fields (cf.~\cite{DBLP:journals/corr/abs-1805-12118}). In place of a metric, the recall in a Top-N set is used as it can be computed for a wide range of algorithms. The dataset is further amended with aggregated statistics about users and items and metadata of an exemplary experiment.

@@

The corpus features comprehensive information on the users~(donors), items~(projects) and transactions respectively interactions~(donations). Transactions are represented by an interaction of a user with an item at a specific point in time and further include a strength~(amount~donated). Users feature information about their location and a boolean value indicating whether they are a teacher. Items feature multiple textual descriptions, funding goals, a list of associated categories and types, information about the creator and details about the location of the school which the item is about.

In a learning subsystem step, two collaborative filtering techniques and two content-based filtering techniques are applied. The performance measures of the single algorithms are amended to the table. Next, aggregated user and item statistics are calculated and added to the table. Lastly, four exemplary meta-learners are applied and the performance measures of the individual approaches are added to the corpus.

@@

The choice of evaluation method ensures that it can be applied to a vast set of different algorithms as long as a ranking of items can be produced. Therefore, it allow for a high variety within the set of algorithms in the learning subsystem in contrast to, e.g., RMSE. It further ensures that not just the ability to decompose the data is measured but the ability to recommend new items to users~\cite{Cremonesi:2010:PRA:1864708.1864721}. Hereby, the learning subsystem is kept extensible while still providing useful metadata to meta-learning algorithms.

@@

The third meta-learner is a classical stacking ensemble using a decision tree which is given the prediction from the learning subsystem as additional input. Based on the metadata and the output of the learning subsystem, it then performs a transaction classification. The final approach aims at solving the ASP via clustering transactions based on the meta-features using K-Means and assigns clusters an overall best algorithm via majority voting.

@@

The requirement enforcing users to have interacted at least twice results in a distribution of the number of users interacting with items which decays exponentially as depicted in~\autoref{fig:InteractionFrequency}. This is to be expected as the time users can invest in researching items and their total interaction strength (budget) is limited. The dataset is dominated by users which interacted exactly two times.

@@

Statistical information about the performances are visualized in~\autoref{fig:MetaLearnerStatistics}. It can be observed that the performance of the learning subsystem depends highly on the employed filtering techniques. Overall, content-based filters are significantly better in achieving a low recall-position (average position in Top-N set; content-based filtering:~$19.48$, collaborative filtering:~$83.28$). The algorithm with the poorest performance is KNN. SVD performs better but is still unable to reliably yield good recommendations. Furthermore, grouping users by shared item interactions has no observable positive impact on the recommendation of new items.

>>>>

\section{Introduction}
Meta-Learning is the process of applying machine learning algorithms on metadata of previous machine learning experiments. The goal is to improve existing approaches by combining the strengths of several single machine learning systems~\cite{AIR:MetalearningSurveyOfTrendsAndTechnologiesLemke2015}. In recommender systems it is an emerging field of interest as meta-learning promises to utilize patterns in metadata as to retrieve the best possible match of item and user. In order to investigate meta-learning algorithms, the required metadata needs to be computed in advance.

One challenge for researchers interested in meta-learning is that suitable datasets are time-consuming and cumbersome to construct. Creating a system of individual machine learning experiments and amending their metadata to the original corpus takes time and resources away from working with the actual meta-learners. Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}. Furthermore, the choice of the metric for computing metadata in the system of single machine learners is important. Using the RMSE, as is done in some of the aforementioned publications, disallows using content-based methods.

Our objective is to address the data augmentation part in a comprehensive and reproducible~\cite{UMUAI:TowardsReproducibilityInRecSysBeel2016} way as to allow researchers to focus on the actual meta-learning task. We are presenting a public domain dataset which to our best knowledge has not been used for meta-learning before in the scientific community, yet due to its feature-rich and vast corpus is a suitable candidate. The raw dataset is augmented with metadata of several machine learning algorithms from a variety of fields (cf.~\cite{DBLP:journals/corr/abs-1805-12118}). In place of a metric, the recall in a Top-N set is used as it can be computed for a wide range of algorithms. The dataset is further amended with aggregated statistics about users and items and metadata of an exemplary experiment.

@@

The corpus features comprehensive information on the users (donors), items (projects) and transactions respectively interactions (donations). Transactions are represented by an interaction of a user with an item at a specific point in time and further include a strength (amount~donated). Users feature information about their location and a boolean value indicating whether they are a teacher. Items feature multiple textual descriptions, funding goals, a list of associated categories and types, information about the creator and details about the location of the school which the item is about.

The learning subsystem comprises two collaborative filtering techniques and two content-based filtering techniques. The performance measures of the single algorithms are amended to the table. Next, aggregated user and item statistics are calculated and added to the table. Lastly, four exemplary meta-learners are applied and the performance measures of the individual approaches are added to the corpus.

@@

The choice of evaluation method ensures that it can be applied to a vast set of different algorithms as long as a ranking of items can be produced. Therefore, it allows for a high variety within the set of algorithms in the learning subsystem in contrast to, e.g., RMSE. It further ensures that not just the ability to decompose the data is measured but the ability to recommend new items to users~\cite{Cremonesi:2010:PRA:1864708.1864721}. Hereby, the learning subsystem is kept extensible while still providing useful metadata to meta-learning algorithms.

@@

The third meta-learner is a classical stacking ensemble using a decision tree which is given the prediction from the learning subsystem as additional input. Based on that, it performs a transaction classification. The final approach aims at solving the ASP via clustering transactions based on the meta-features using K-Means and assigns clusters an overall best algorithm via majority voting.

@@

Requiring users to have interacted at least twice, results in a distribution of the number of interactions per user which decays exponentially as depicted in~\autoref{fig:InteractionFrequency}. This is to be expected as the time and interaction strength (budget) a user can invest is limited. The dataset is dominated by users which interacted exactly two times.

@@

Statistical information about the performances are visualized in~\autoref{fig:MetaLearnerStatistics}. It can be observed that the performance of the learning subsystem depends highly on the employed filtering techniques. Overall, content-based filters are significantly better in achieving a low recall-position (average position in Top-N set; content-based filtering:~$19.48$, collaborative filtering:~$83.28$). The algorithm with the poorest performance is KNN. SVD performs better but is still unable to reliably yield good recommendations. Grouping users by shared item interactions has no observable positive impact on the recommendation.
